{RandomRationalPoints} package for \emph{Macaulay2}

### 2020-12-21 FastLinAlg/Paper/RandomRationalPoints.tex
Some Further Changes,Example 3.4 added

 We provide the following core functions:
 \begin{itemize}
	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
 	\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
    \item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by findANonZeroMinor. (\Cref{extendIdealByNonZeroMinor})	
 \end{itemize}

@@

%	\item {\tt Codimension}: It takes input an integer. Various functions need to know the codimension/height of the scheme/ideal it is working with.  Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.  The default value is {\tt null}, in which case the function will compute the codimension.  
%%%
%%Codimension is not a function, it is an option.
%%%
\subsection*{\tt randomCoordinateChange} This function takes a polynomial ring as an input and outputs the coordinate change map.  
	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

	\begin{itemize}
		\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.  If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications. If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
		\item {\tt MaxCoordinatesToReplace}: the user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
		This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
		\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
	\end{itemize}

@@

	o7 = | y x+4z z |
	\end{verbatim}}
	}

\section{genericProjection, projectionToHypersurface}\label{projectionfunctions}

@@

%\begin{enumerate}
\subsection{genericProjection} 
This function finds a random (somewhat depending on options) generic projection of the ring or ideal.
The typical usages are as follows: 
\vspace{1em}

@@

\vspace{1em}
\noindent where 
$I$ is an ideal 
in a polynomial ring, 
$R$ can denote a quotient of a polynomial ring and 
$n\in \mathbb{Z}$ is
an integer specifying how many dimensions to drop.  Note, this function makes no attempt to verify that the projection is actually generic with respect to the ideal.

@@

Alternately, instead of {\tt I}, we may pass it a quotient ring.  It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.  It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.

@@

@@

@@

\vspace{1em}
\noindent where $n$ is a positive integer denoting the number of desired points, and 
$I$ is an ideal inside a polynomial ring.  If {\tt n} is omitted, then it is assumed to be 1.

@@

	Notice that the speed, or success, varies depending on the strategy.

@@

	\subsection{\tt findANonZeroMinor}\label{findANonZeroMinor}: The typical usage of this function is as follows:

@@

\begin{description}
	\item[\tt Strategy => Symbol] to specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).

@@

\subsection{\tt extendIdealByNonZeroMinor}\label{extendIdealByNonZeroMinor}: The typical usage is

@@

In this particular example, there tend to be about 5 associated primes when adding the first minor to J, and so one would expect about 5 steps as each minor computed most likely will eliminate one of those primes.

>>>>

 We provide the following core functions:
 \begin{itemize}
		\item {\tt projectionToHypersurface} and {\tt genericProjection}: These functions provide customizable projection. (\Cref{projectionfunctions}) 	
 	\item {\tt randomPoints}:  This tries to find a point in the vanishing set of an ideal. (\Cref{randomPoints})
    \item {\tt findANonZeroMinor}:  This finds a submatrix of a given matrix that is nonsingular at a point of a given ideal. (\Cref{findANonZeroMinor})          
	\item {\tt extendIdealByNonZeroMinor}:  This extends an ideal by a minor produced by \\{\tt findANonZeroMinor}. (\Cref{extendIdealByNonZeroMinor})	
 \end{itemize}

@@

%	\item {\tt Codimension}: It takes input an integer. Various functions need to know the codimension/height of the scheme/ideal it is working with.  Setting this to be an integer will tell the function not to compute the codimension and to use this value instead.  The default value is {\tt null}, in which case the function will compute the codimension.  
%%%
%%Codimension is not a function, it is an option.
%%%
\subsection*{\tt randomCoordinateChange:} This function takes a polynomial ring as an input and outputs the coordinate change map.  
	i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

	\begin{itemize}
		\item {\tt Replacement}: Setting {\tt Replacement => Full} will mean that coordinates are replaced by a general degree $1$ form.  If {\tt Replacement => Binomial}, the coordinates are only changed to binomials, which can be much faster for certain applications. If {\tt Homogeneous => false}, then there will be constant terms, and we view $mx + b$ as a binomial.
		\item {\tt MaxCoordinatesToReplace}: The user can specify that only a specified number of coordinates should be non-monomial (assuming {\tt Homogeneous => true}).  
		This option is passed to {\tt randomCoordinateChange} by other functions that call it. 		
		\item {\tt Homogeneous}:  Setting {\tt Homogeneous => false} will cause degree zero terms to be added to modified coordinates (including monomial coordinates).		
	\end{itemize}

@@

	o7 = | y x+4z z |
	\end{verbatim}}
	}}

\end{example}
\section{genericProjection, projectionToHypersurface}\label{projectionfunctions}

@@

%\begin{enumerate}
\subsection{genericProjection} 
This function finds a random (somewhat, depending on options) generic projection of the ring or ideal.
The typical usages are as follows: 
\vspace{1em}

@@

\vspace{1em}
\noindent where 
$I$ is an ideal 
in a polynomial ring, 
$R$ can denote a quotient of a polynomial ring and 
$n\in \mathbb{Z}$ is
an integer specifying how many dimensions to drop.  Note that this function makes no attempt to verify that the projection is actually generic with respect to the ideal.

@@

Alternately, instead of {$I$}, we may pass it a quotient ring.  It will then return the inclusion of the generic projection ring into the given ring, followed by the source of that inclusion.  It is essentially the same functionality as calling {\tt genericProjection(n, ideal R)} although the format of the output is slightly different.

@@

\newpage

@@

\vspace{1em}
\begin{example}
	{{\small\color{blue}
	\begin{verbatim}
	i2 : R=ZZ/5[x,y,z];

	               3       2     3            2        2       2       2  
	o3 = ideal (- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  -
	-------------------------------------------------------------------
	        2
	2x*y - y  + x*z - y*z)

	i4 : projectionToHypersurface(I)

	            ZZ                             4    3      2 2     4
	o4 = (map(R,--[y, z],{x + 2y, 2y}), ideal(y  + y z - 2y z  - 2z  - y -
	            5
	-------------------------------------------------------------------
	z))

	o4 : Sequence

	i5 : projectionToHypersurface(R/I)

                           	                    R                    
	o5 = (map(--------------------------------------------------------------
	    3       2     3            2        2       2       2     
	(- x  - 2x*y  - 2y  - x*y*z + y z - 2x*z  + 2y*z  - 2, x  - 2x
	-------------------------------------------------------------------
                                    ZZ                       
	                                    --[y, z]                 
	                                    5                       
	--------------------,----------------------------------------------
	      2               4 2    3 3       5     6     3       2     3 
	*y - y  + x*z - y*z) y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  
	-------------------------------------------------------------------
                    	       ZZ
                	           --[y, z]
	                            5
	---,{x - z, y}), -------------------------------------------------)
	                  4 2    3 3       5     6     3       2     3
	- 1              y z  + y z  - 2y*z  + 2z  + 2y  + 2y*z  + 2z  - 1

	o5 : Sequence

@@

\vspace{1em}
\noindent where $n$ is a positive integer denoting the number of desired points, and 
$I$ is an ideal inside a polynomial ring.  If {$ n$} is omitted, then it is assumed to be 1.

@@

	Notice that the speed, or success, varies depending on the strategy.
\newpage

@@

	\subsection{\tt findANonZeroMinor:}\label{findANonZeroMinor} The typical usage of this function is as follows:

@@

\begin{description}
	\item[\tt Strategy => Symbol] To specify which strategy to use when calling {\tt randomPoints} (see \Cref{strategydetails}).

@@

\subsection{\tt extendIdealByNonZeroMinor:}\label{extendIdealByNonZeroMinor} The typical usage is

@@

In this particular example, there tend to be about $5$ associated primes when adding the first minor to $J$, and so one would expect about $5$ steps as each minor computed most likely will eliminate one of those primes.
